# Section 00 - Getting ready
1. Make sure you have already installed and loaded the following libraries:
```{r}
library(ggplot2)
library(data.table)
library(magrittr) # Needed for %>% operator
library(tidyr)
library(ggrepel)
```

# Section 01 - Plot types
Match each chart type with the relationship it shows best.
1. shows distribution and quantiles, especially useful when comparing uni-modal distributions.
boxplot

2. highlights individual values, supports comparison and can show rankings or deviations categories and totals
bar chart

3. shows overall changes and patterns, usually over intervals of time
line chart

4. shows relationship between two continuous variables.
scatterplot

Options: bar chart, line chart, scatterplot, boxplot

# Section 02 - Reproducing a plot with ggplot2
1. Reproduce the following visualization of the association between the variables cty and hwy for the years
1999 and 2008 from the dataset mpg using the library ggplot2:
```{r}
data(mpg)
mpg <- as.data.table(mpg)

mpg[, year:=as.factor(year)]

ggplot(data=mpg, aes(x=cty, y=hwy, color=year)) +
  geom_point() +
  geom_smooth(method="lm")
```

