# Section 00 - Getting ready
1. Make sure you have already installed and loaded the following libraries:
```{r}
library(ggplot2)
library(data.table)
library(magrittr) # Needed for %>% operator
library(tidyr)
```

# Section 01 - Color guidelines
What are best practices when using color for data visualizations? Select all that apply.
1. Avoid having too many colors for categorical data. x
2. Use color only when it actually adds meaning to the plot. x
3. Use divergent color scales for categorical data types.

# Section 02 - Correlation and Causation
Read the following statements. Decide for each statement separately whether it is true or false and give an
explanation.
1. The concept of reverse causality states that whenever A causes B, B also causes A. FALSE
2. If A causes B and A causes C, then B also causes C. FALSE
3. If A and B correlate and A happens before B, then A causes B. FALSE
4. Causation implies linear association. FALSE

# Section 03 - Effect of a third variable
Investigate the file coffee_sim.csv (simulated dataset) by first loading it as a data.table.
```{r}
coffee_dt <- fread("./extdata/coffee_sim.csv")
coffee_dt
summary(coffee_dt)
```
1. Suggest an appropriate visualization and implement it with ggplot2 to display a possible association
between coffee consumption and “datavizitis” disease risk, measured in deaths per 1000 individuals. Does
this plot by itself seem consistent with a causal effect of coffee on datavizitis?
```{r}
coffee_dt[,packs_cigarettes_per_day := as.factor(packs_cigarettes_per_day)]
coffee_dt[,coffee_cups_per_day := as.factor(coffee_cups_per_day)]

ggplot(coffee_dt, aes(coffee_cups_per_day, datavizitis_risk)) +
  geom_boxplot()

```
2. Investigate the full dataset. Do you see evidence for a third variable influencing association? Support your
statement with an appropriate plot. Draw a graph with the potential causal relationships you find consistent
with the data. Relate it to one of the situations from the lecture script’s figure 6.3 or Simpson’s paradox.
```{r}
ggplot(coffee_dt, aes(packs_cigarettes_per_day, datavizitis_risk)) +
  geom_boxplot()

ggplot(coffee_dt, aes(coffee_cups_per_day, datavizitis_risk, color=packs_cigarettes_per_day)) +
  geom_boxplot()


ggplot(coffee_dt, aes(packs_cigarettes_per_day, datavizitis_risk, color=coffee_cups_per_day)) +
  geom_boxplot()


ggplot(coffee_dt, aes(coffee_cups_per_day, datavizitis_risk)) +
  geom_boxplot() +
  facet_wrap(~packs_cigarettes_per_day)


ggplot(coffee_dt, aes(packs_cigarettes_per_day, datavizitis_risk)) +
  geom_boxplot() +
  facet_wrap(~coffee_cups_per_day)

```
common cause
smoking -> datavisitis
smoking -> coffee


# Section 04 - General guidelines in data visualization
Below is a graph taken from one published paper. Read the figure legend.
