---
title: "R Notebook"
output: html_notebook
---

```{r}
library(data.table)
library(magrittr)
```
# Section 1
1
```{r}
ratings_dt <- fread("./extdata/Lecture02_extdata/Exercise_data/BX-Book-Ratings.csv")
books_dt <- fread("./extdata/Lecture02_extdata/Exercise_data/BX-Books.csv")
users_dt <- fread("./extdata/Lecture02_extdata/Exercise_data/BX-Users.csv")
```

3
```{r}
users_dt$Age <- as.integer(users_dt[, Age])
users_dt$Age[is.na(users_dt$Age)] <- 0

# Alternative
# users_dt[, Age := as.numeric(Age)]

```
4 5
```{r}
summary(books_dt)
sapply(users_dt, class)

head(ratings_dt, n=5)
tail(ratings_dt, n=5)
```
6
```{r}
colnames(ratings_dt) <- gsub('-', '_', colnames(ratings_dt))  
colnames(books_dt) <- gsub('-', '_', colnames(books_dt))  
colnames(users_dt) <- gsub('-', '_', colnames(users_dt))  
```

7
```{r}
books_dt[, c("Image_URL_S", "Image_URL_M", "Image_URL_L") := NULL]
```

8
Create a table book_dt_2 that contains all the books published between 1900 and 2019 (inclusive) from
the table books_dt.
```{r}
book_dt_2 <- books_dt[Year_Of_Publication >= 1900 & Year_Of_Publication <= 2019]
```

# Section 2
1
```{r}
length(unique(books_dt[, Book_Author]))

# Alternative
books_dt[, uniqueN(Book_Author)]
```

2
How many different authors are included for each year of publication between 2000 and 2010 (inclusive)
in books_dt?
```{r}
books_dt[Year_Of_Publication >= 2000 & Year_Of_Publication <= 2010, uniqueN(Book_Author), by=Year_Of_Publication]
```

3
In how many observations is the age information missing in the ratings table users_dt?
```{r}
users_dt[Age == 0, .N]
```

4
What is the maximum rating value in the ratings table?
```{r}
max(unique(ratings_dt[, Book_Rating]))

ratings_dt[, max(Book_Rating, rm.na=TRUE)]
```

5. What is the most common rating value larger than 0?
```{r}
ratings_dt[Book_Rating != 0, .N, by = Book_Rating][N==max(N)]
```

6. Which are the book identifiers (ISBN) with the highest ratings?
```{r}
ratings_dt[Book_Rating == 10, ISBN] %>% head

ratings_dt[Book_Rating == max(Book_Rating, rm.na=TRUE), ISBN] %>% head
```

7. Reorder the ratings table according to the rating value of each book in descending order. Hint: order()
```{r}
ratings_dt[order(Book_Rating, decreasing = FALSE)]

ratings_dt[order(-Book_Rating)]

# setorder(ratings_dt, -Book_Rating)
```

# Section 3
1. Add a new column called High_Rating to the data table ratings_dt. The column has an integer 1 for
all observations with a rating value higher than 7. Hint: ifelse()
```{r}

```

2. How many observations are considered to be a high ranking? What is the proportion of high ranked
observations among all observations?
3. Which users did not give any rating to any book? Filter these users out from users_dt. Hint: There’s no
need to merge users_dt with ratings_dt, we are simply interested in the users that are not in ratings_dt.
4. What is the most common age of users who rated at least one book?
5. On average, how many books did a user rate?
6. What is the title of the first published book with the highest ranking?
7. In which year was a book with the largest number of ratings last published?
8. Add to the table ratings_dt the highest ranking that each book received as a new column called
Max_Book_Ranking.
9. Subset the ratings_dt ratings table to contain only books written by the following authors:






