---
title: "R Notebook"
output: html_notebook
---


```{r}
library(data.table)
library(magrittr)
library(tidyr)
```

# Section 1

Hint: You can solve parts 1, 2 and 3 by looking at the printed output. You do not need to load the datasets.

1. Visually examine the dataset AirPassengers. Which of the following is true:
a. AirPassengers is tidy data: it has one year for each row.
b. AirPassengers is not tidy: we need at least one column with a character vector.
c. AirPassengers is not tidy: it is a matrix instead of a data frame.
-> d. AirPassengers is not tidy: to be tidy we would have to wrangle it to have three columns (year, month
and value), then each passenger count would have a row.

2. Visually examine the dataset ChickWeight. Which of the following is true:
a. ChickWeight is not tidy: each chick has more than one row.
-> b. ChickWeight is tidy: each observation (a weight) is represented by one row. The chick from which this
measurement came is one of the variables.
c. ChickWeight is not tidy: we are missing the year column.
d. ChickWeight is tidy: it is stored in a data frame.

3. Visually examine the dataset spanish_vowels. Is the data set tidy?
No Label can be split

4. The example_product_data.csv file describes the number of times a person bought product “a” and “b”.
Load the file into a data.table.
```{r}
product_dt <- fread("./extdata/example_product_data.csv")
product_dt
```


5. Transform product_dt into a long format using data.table commands.
```{r}
long_dt <- melt(product_dt, id.vars = "name", measure.vars = c("producta", "productb"), variable.name = "product", value.name = "count")
long_dt <- melt(product_dt, id.vars = "name", variable.name = "product", value.name = "count")
long_dt
```

6. Transform the table from the long format back into a wide format. Check that it is equal to the original
data.table.
```{r}
dcast(long_dt, ... ~ product, value.var = "count")
```

# Section 02 - Merge Warm Up

Prepare two tables by running the following code:
```{r}
mtcars_dt <- as.data.table(mtcars)
mtcars_dt[, carname := rownames(mtcars)]
dt1 <- mtcars_dt[5:25,.(carname, mpg, cyl)]
dt2 <- mtcars_dt[1:10, .(carname, gear)]

dt1
dt2
```

1. How many rows does the inner merge of dt1 and dt2 have? 6
2. How many rows does the left merge of dt1 and dt2 have? 21
3. How many rows does the outer merge of dt1 and dt2 have? 25

```{r}
merge(dt1, dt2, by = "carname")
# or all = FALSE

merge(dt1, dt2, by = "carname", all.x = TRUE)

merge(dt1, dt2, by = "carname", all = TRUE)
```

# Section 03 - Weather dataset

1. Read in the weather dataset weather.txt as a data.table. Please find this file on Moodle and save it to
a folder called extdata in your working directory.
```{r}
weather_dt <- fread("./extdata/weather.txt")
weather_dt
```

2. Why is this dataset messy?
Different days are seperate col names
Also many NA values
3. How would a tidy version of it look like? Do not give the code, only describe how the tidy table would
look like.
Add another col called day and have it contain the value of the (1-31)

4. Create a tidy version of the weather dataset.
```{r}
better_weather_dt <- na.omit(melt(weather_dt, id.vars = c("id", "year", "month", "element"), variable.name = "day", value.name = "temperature"))
better_weather_dt

better_weather_dt[, day:=substr(day, 2, length(day))]
better_weather_dt

better_weather_dt <- dcast(better_weather_dt, ... ~ element, value.var = "temperature")
better_weather_dt

better_weather_dt <- better_weather_dt[, .(id, day, month, year, TMAX, TMIN)]
better_weather_dt

# I feel this is not needed but it is done in the example
better_weather_dt <- better_weather_dt[, .(id, date=paste(day, month, year, sep = "/"), TMAX, TMIN)]
better_weather_dt
```

# Section 04 -Scattered data across many files

The baby-names folder contains 258 csv-files (1999.girl.csv, 1999.boy.csv , . . . ) which store name
frequencies for a particular year and sex.

1. Create a list containing all file paths in the folder, i.e. (extdata/baby-names/1999.girl.csv,
extdata/baby-names/1999.boys.csv, . . . ). Refer to Appendix A of the script if you are having trouble
with the file paths.
```{r}
paths <- list()

for(year in 1880:2008) {
  paths <- append(paths, file.path("extdata", "baby-names", paste(year, ".boy.csv", sep = "") ))
  paths <- append(paths, file.path("extdata", "baby-names", paste(year, ".girl.csv", sep = "") ))
}

# Better way of doing this
paths <- list.files("extdata/baby-names", full.names=TRUE)
paths
```

2. Name the list entries with the basename of the corresponding file path. The basename of a path is the
actual name of the file, e.g. the basename of extdata/baby-names/1999.girl.csv is 1999.girl.csv.
```{r}
names(paths) <- basename(paths)
paths
```

3. Read in the data from all files into one table. Hint: when you read many files and gather them into one
table, be sure to add a column that identifies each file. rbindlist()
```{r}
tables <- lapply(paths, fread)

baby_names_dt <- rbindlist(tables, use.names = TRUE, idcol = "file_name")
baby_names_dt
```

4. Is the data tidy? If not, tidy it up.
No, gender and year can be split
```{r}
baby_names_dt <- as.data.table(separate(baby_names_dt, col = "file_name", into = c("year", "gender"), extra = "drop"))
baby_names_dt
baby_names_dt[, gender:=ifelse(gender=="boy", "M", "F") ]
baby_names_dt
```
# Section 05 - Small case-study: cleaning up a gene-expression dataset in yeast

In this section, we will read and clean up the data from the paper:
• Gagneur, Julien, et al. “Genotype-environment interactions reveal causal pathways that mediate genetic
effects on phenotype.” PLoS Genet 9.9 (2013): e1003803. https://journals.plos.org/plosgenetics/artic
le?id=10.1371/journal.pgen.1003803

You will not need to understand the data in depth for this exercise but if you are interested you can find a
detailed description in section 7.1 of the script.

1. Read in the two files in the folder eqtl.
The first file contains the genotypes of yeast strains and a strain identifyer. The second file contains
information on how quickly each strain grows in different growth media.
```{r}
gt <- fread('extdata/eqtl/genotype.txt')
dim(gt)
head(gt[,1:5])
gt
```
```{r}
growth <- fread('extdata/eqtl/growth.txt')
growth
```

2. Come up with a strategy, how you can transform the two tables shown above into the single table shown
below.
head(dt)
## strain media growth_rate marker gt
## 1: seg_01B YPD 12.60399 mrk_1 Lab strain
## 2: seg_01B YPD 12.60399 mrk_14 Lab strain
## 3: seg_01B YPD 12.60399 mrk_27 Lab strain
## 4: seg_01B YPD 12.60399 mrk_40 Lab strain
## 5: seg_01B YPD 12.60399 mrk_54 Lab strain
## 6: seg_01B YPD 12.60399 mrk_67 Lab strain
summary(dt)
## strain media growth_rate marker
## seg_01B: 5000 YPD :158000 Min. : 1.57 mrk_1 : 790
## seg_01C: 5000 YPD_BPS :158000 1st Qu.: 4.55 mrk_14 : 790
## seg_01D: 5000 YPD_Rapa:158000 Median : 6.93 mrk_27 : 790
## seg_02B: 5000 YPE :158000 Mean : 7.60 mrk_40 : 790
## seg_02C: 5000 YPMalt :158000 3rd Qu.:10.70 mrk_54 : 790
## seg_02D: 5000 Max. :16.27 mrk_67 : 790
## (Other):760000 NA's :42000 (Other):785260
## gt
## Lab strain :398145
## Wild isolate:391855

```{r}
growth_long_dt <- melt(growth, id.vars = c("strain"), variable.name = "media", value.name = "growth_rate")
growth_long_dt

gt_long_dt <- melt(gt, id.vars = c("strain"), variable.name = "marker", value.name = "gt")
gt_long_dt

merged_dt <- merge(growth_long_dt, gt_long_dt, by = "strain", allow.cartesian = TRUE)
merged_dt

merged_dt[,strain:=as.factor(strain)]
merged_dt[,gt:=as.factor(gt)]
merged_dt

summary(merged_dt)
```
3. Write code that implements your strategy to transform the two tables into the one shown above. (see above)
4. Spoiler: These are the things you can do with tidy data. When you are done run the following line of code
and observe the result:
```{r}
library(ggplot2)

dt <- merged_dt
ggplot(dt[marker %in% c('mrk_5211', 'mrk_1653')], aes(marker, growth_rate, color=gt)) +
geom_boxplot() + facet_wrap(~media)
```