Section 00 - Getting ready
1. Make sure you have already installed and loaded the following libraries:
```{r}
library(ggplot2)
library(data.table)
library(magrittr) # Needed for %>% operator
library(tidyr)
library(GGally)
library(pheatmap)
library(mclust)
```


# Section 01 - Visualizing multiple variables
In this exercise, we will revisit how to do a correlation analysis and potential pitfalls of correlation analysis.
We will use the gene expression data in cancer_data.rds which stores the expression of 20 genes across 30
different tumor samples. Gene expression measures the abundance of RNAs per gene and is indicative of how
active a gene is in a sample. If the expression of a gene differs a lot between conditions (e.g. healthy versus
cancer or different cancer types) it could hint that this gene plays an important role in this context.
Load the gene expression data in cancer_data.rds as a data.table with the following line of code:
```{r}
expr <- readRDS("extdata/cancer_data.rds") %>% as.data.table(keep.rownames="tumor_type")
expr
```
## tumor_type MYC SRM GBE1 FUK UGP2
## 1: DOHH2 -0.4950154 0.20907327 -0.4366726 100.0000000 100.0000000
## 2: FARAGE -0.4913630 -0.35590874 0.5322076 0.1037538 -0.6494896
## 3: HT -0.4559935 -0.23852493 0.6714241 -1.0136358 -0.9651291
## 4: Kapas231 1.6152666 0.92666297 1.0168147 -0.9978958 1.0064845
## 5: OCI-LY1 0.4942733 2.40062582 -1.4255166 1.0676067 0.9862178
## 6: OCI-LY1-B50 0.1940777 -0.04941387 -0.4810955 0.2423189 0.7046970

1. We are interested in the correlations between genes. Plot the pairwise correlations of the variables in the
dataset. Which pair of genes has the highest correlation? Hint: remember that you can exclude a column
"colA" from a data table DT with DT[, -"colA"].
```{r}
gene_cor <- cor(expr[, -"tumor_type"])
pheatmap(gene_cor)

# alternative
ggcorr(expr[, -"tumor_type"]) 
```

2. Visualize the raw data in a heatmap with pheatmap. Hint: pheatmap does not work well with data.tables,
you should therefore convert it to a matrix before plotting with as.matrix()
```{r}
expr_mat <- as.matrix(expr[, -"tumor_type"])
rownames(expr_mat) <- expr[, tumor_type]

pheatmap(expr_mat)
```

3. Does the latter plot suggest some outliers? Could they have affected the correlations? Check by using an
appropriate plot the impact of these outliers on the correlations in question 1. Substitute them with missing
values (NA) and redo the previous questions 1 and 2.
```{r}
expr_long <- melt(expr, id.vars = "tumor_type", variable.name = "gene", value.name = "expression")
expr_long[order(-expression),]

ggplot(expr, aes(FUK, UGP2)) +
  geom_point()

expr[tumor_type=="DOHH2", "FUK"] <- NA
expr[tumor_type=="DOHH2", "UGP2"] <- NA
expr
```

```{r}
gene_cor <- cor(expr[, -"tumor_type"], use="complete.obs")
pheatmap(gene_cor)

# alternative
ggcorr(expr[, -"tumor_type"]) 


expr_mat <- as.matrix(expr[, -"tumor_type"])
rownames(expr_mat) <- expr[, tumor_type]

pheatmap(expr_mat)
```

# Section 02 - Heatmaps and Hierarchical clustering
1. Consider the full iris data set without the Species column for clustering. Create a pretty heatmap with
the library pheatmap of the data without clustering.
```{r}
data(iris)

iris
iris$Species = NULL

iris_mat <- as.matrix(iris)

pheatmap(iris_mat, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column", labels_row = 0)
```

2. Now, create a pretty heatmap using complete linkage clustering of the rows of the data set. Hint: You
can specify a clustering method with the clustering_method argument in pheatmap
```{r}
pheatmap(iris_mat, cluster_rows = TRUE, cluster_cols = TRUE, clustering_method = "complete", scale = "column")
```

3. Obtain the dendrogram of the row clustering using complete linkage clustering and partition the data
into 3 clusters. Hint: You can use cutree for cutting the tree.
```{r}
heatmap <- pheatmap(iris_mat, cluster_rows = TRUE, cluster_cols = TRUE, clustering_method = "complete", scale = "column")

trees <- cutree(heatmap$tree_row, k = 3)
```

4. Annotate the rows of the heatmap with the Species column of the iris dataset and the three clusters
from complete linkage clustering. What do you observe when you compare the clustering and the species
labels?
```{r}
data(iris)

annotation <- data.frame( complete = paste0("T",trees),
                       species = iris$Species)
rownames(annotation) <- rownames(iris)

rownames(iris_mat) <- rownames(iris)

pheatmap(iris_mat, cluster_rows = T, cluster_cols = T, clustering_method = "complete", scale = "column", annotation_row = annotation, show_rownames = F)
```










