Section 00 - Getting Ready
1. Make sure you have already installed and loaded the following libraries:
```{r}
library(ggplot2)
library(data.table)
library(magrittr)
library(tidyr)
library(dplyr)
library(datasets)
```
2. Load yeast data and required packages using the following code:
```{r}
gene <- fread("./extdata/eqtl/gene.txt")
genotype <- fread("./extdata/eqtl/genotype.txt")
genotype <- melt(genotype, id.vars = 'strain', variable.name = 'marker',
value.name = 'genotype')
growth <- fread("./extdata/eqtl/growth.txt")
growth <- melt(growth, id.vars = "strain", variable.name = 'media',
value.name = 'growth_rate')
marker <- fread("./extdata/eqtl/marker.txt")
```

# Section 01 - Warm Up: Choosing the right test
1. You are a data science consultant helping researchers pick the right tests to evaluate their hypotheses. For
each hypothesis, indicate which test from the ones you have seen in the lecture would be most appropriate:

a) A researcher collects data on the height (measured in cm) and weight (measured in g) of Germans.
She hypothesizes that there is a significant association between how tall Germans are and how much
they weigh. She would like to test this hypothesis without making any distributional assumptions.
spearman correlation

b) A researcher collects data on the weight (measured in g) of Bavarians before and after the Oktoberfest.
She would like to know whether there is a significant difference in average weight after the Oktoberfest
as compared to before it. Prior research indicates that the weight of Bavarians is approximately
normally distributed.
t-test (/welch or wilcox)

c) A researcher is evaluating a rapid antigen test. The company manufacturing the test claims that if
someone is infected, the test will correctly return a positive result 99% of the time. The researcher
hypothesizes that, in practice, the test is often improperly administered and therefore significantly less
sensitive. She asks 1000 individuals, which have all been confirmed to be infected by a PCR test, to
self-administer the antigen test. She records how often the antigen test correctly returns a positive
result.
binomial test p=0.99 as null hypothesis

d) The company manufacturing the test has collected a bigger dataset, comprising both infected and
non-infected individuals. For each individual, they record two datapoints: the result of a PCR test
(infected/not-infected), which is taken as ground truth, and the result of a self-administered antigen
test (positive/negative). They would like to show that, even if self-administered, the test still gives
some information about infection status and thus is better than nothing.
fischer exact test
