Section 00 - Getting ready
1. Make sure you have already installed and loaded the following libraries:
```{r}
library(ggplot2)
library(data.table)
library(magrittr) # Needed for %>% operator
library(tidyr)
library(GGally)
library(pheatmap)
library(mclust)
```


# Section 01 - Visualizing multiple variables
In this exercise, we will revisit how to do a correlation analysis and potential pitfalls of correlation analysis.
We will use the gene expression data in cancer_data.rds which stores the expression of 20 genes across 30
different tumor samples. Gene expression measures the abundance of RNAs per gene and is indicative of how
active a gene is in a sample. If the expression of a gene differs a lot between conditions (e.g. healthy versus
cancer or different cancer types) it could hint that this gene plays an important role in this context.
Load the gene expression data in cancer_data.rds as a data.table with the following line of code:
```{r}
expr <- readRDS("extdata/cancer_data.rds") %>% as.data.table(keep.rownames="tumor_type")
expr
```
## tumor_type MYC SRM GBE1 FUK UGP2
## 1: DOHH2 -0.4950154 0.20907327 -0.4366726 100.0000000 100.0000000
## 2: FARAGE -0.4913630 -0.35590874 0.5322076 0.1037538 -0.6494896
## 3: HT -0.4559935 -0.23852493 0.6714241 -1.0136358 -0.9651291
## 4: Kapas231 1.6152666 0.92666297 1.0168147 -0.9978958 1.0064845
## 5: OCI-LY1 0.4942733 2.40062582 -1.4255166 1.0676067 0.9862178
## 6: OCI-LY1-B50 0.1940777 -0.04941387 -0.4810955 0.2423189 0.7046970

1. We are interested in the correlations between genes. Plot the pairwise correlations of the variables in the
dataset. Which pair of genes has the highest correlation? Hint: remember that you can exclude a column
"colA" from a data table DT with DT[, -"colA"].
```{r}
gene_cor <- cor(expr[, -"tumor_type"])
pheatmap(gene_cor)

# alternative
ggcorr(expr[, -"tumor_type"]) 
```

2. Visualize the raw data in a heatmap with pheatmap. Hint: pheatmap does not work well with data.tables,
you should therefore convert it to a matrix before plotting with as.matrix()
```{r}
expr_mat <- as.matrix(expr[, -"tumor_type"])
rownames(expr_mat) <- expr[, tumor_type]

pheatmap(expr_mat)
```

3. Does the latter plot suggest some outliers? Could they have affected the correlations? Check by using an
appropriate plot the impact of these outliers on the correlations in question 1. Substitute them with missing
values (NA) and redo the previous questions 1 and 2.
```{r}
expr_long <- melt(expr, id.vars = "tumor_type", variable.name = "gene", value.name = "expression")
expr_long[order(-expression),]

ggplot(expr, aes(FUK, UGP2)) +
  geom_point()

expr[tumor_type=="DOHH2", "FUK"] <- NA
expr[tumor_type=="DOHH2", "UGP2"] <- NA
expr
```

```{r}
gene_cor <- cor(expr[, -"tumor_type"], use="complete.obs")
pheatmap(gene_cor)

# alternative
ggcorr(expr[, -"tumor_type"]) 


expr_mat <- as.matrix(expr[, -"tumor_type"])
rownames(expr_mat) <- expr[, tumor_type]

pheatmap(expr_mat)
```

# Section 02 - Heatmaps and Hierarchical clustering
1. Consider the full iris data set without the Species column for clustering. Create a pretty heatmap with
the library pheatmap of the data without clustering.
```{r}
data(iris)

iris
iris$Species = NULL

iris_mat <- as.matrix(iris)

pheatmap(iris_mat, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column", labels_row = 0)
```

2. Now, create a pretty heatmap using complete linkage clustering of the rows of the data set. Hint: You
can specify a clustering method with the clustering_method argument in pheatmap
```{r}
pheatmap(iris_mat, cluster_rows = TRUE, cluster_cols = TRUE, clustering_method = "complete", scale = "column")
```

3. Obtain the dendrogram of the row clustering using complete linkage clustering and partition the data
into 3 clusters. Hint: You can use cutree for cutting the tree.
```{r}
heatmap <- pheatmap(iris_mat, cluster_rows = TRUE, cluster_cols = TRUE, clustering_method = "complete", scale = "column")

trees <- cutree(heatmap$tree_row, k = 3)
```

4. Annotate the rows of the heatmap with the Species column of the iris dataset and the three clusters
from complete linkage clustering. What do you observe when you compare the clustering and the species
labels?
```{r}
data(iris)

annotation <- data.frame( complete = paste0("T",trees),
                       species = iris$Species)
rownames(annotation) <- rownames(iris)

rownames(iris_mat) <- rownames(iris)

pheatmap(iris_mat, cluster_rows = T, cluster_cols = T, clustering_method = "complete", scale = "column", annotation_row = annotation, show_rownames = F)
```

# Section 03 - Cluster comparison
1. Compute the Rand index between the two following clustering results from two different clustering
algorithms. You can solve this exercise with pen and paper.

11/15 ~= 0.73

# Section 04 - Clustering and Heatmaps
In this section, we aim to compare two different clustering algorithms - hierarchical clustering and k-means
clustering.
1. This plot represents a random initialization of a k-means algorithm with k=2. X1, X2 are the randomly
positioned centroids and A to E are the points of the 2-dimensional dataset. Calculate the new positions
of the centroids after the first iteration using the euclidean distance. Hint: You can solve this exercise
either with pen and paper or using R.
```{r}
# X1 6 9
# X2 3 3

points <- data.frame(x = c(08,10,04,05,06),
                     y = c(05,07,06,08,04))
rownames(points) <- c("A", "B", "C", "D", "E")
points

points <- as.data.table(points)

points[, dist_x1 := sqrt((x-6)^2 + (y-9)^2)]
points[, dist_x2 := sqrt((x-3)^2 + (y-3)^2)]
points[, closer := ifelse(dist_x1 < dist_x2, "new_x1", "new_x2")]
points

points[, .(x=mean(x), y=mean(y)), by=closer]

ggplot(points, aes(x,y, label=rownames(points))) + 
  geom_point() +
  geom_label()

```


2. Perform k-means clustering on the iris data set with k = 3.
```{r}
data(iris)
iris_mat <- as.matrix(iris[, 1:4])

iris_mat_scaled <- scale(iris_mat)

kmeans_cluster <- kmeans(iris_mat_scaled, 3, nstart=100) 
kmeans_cluster$cluster
```

3. Reproduce section 2 from the tutorial. Use the table function to compare the partitions from the complete
linkage and k-means clustering.
```{r}
heatmap <- pheatmap(iris_mat, cluster_rows = TRUE, cluster_cols = TRUE, clustering_method = "complete", scale = "column")
complete_link_clusters <- cutree(heatmap$tree_row, k = 3)

table(complete_link_clusters, kmeans_cluster$cluster)
```

4. Create a pretty heatmap using complete clustering of the rows annotated with the species and both
clustering results - complete linkage clustering and the k-means clustering. What do you observe when you
compare the two different clustering algorithms and the species labels?
```{r}
data(iris)

annot <- data.frame(species = iris$Species,
                    complete_link = complete_link_clusters,
                    kmeans = kmeans_cluster$cluster)
rownames(annot) <- rownames(iris)

rownames(iris_mat) <- rownames(iris)

pheatmap(iris_mat, cluster_rows = T, cluster_cols = T, clustering_method = "complete", scale = "column", annotation_row = annot, show_rownames = F)
```









