Section 00 - Getting Ready
1. Make sure you have already installed and loaded the following libraries:
```{r}
library(ggplot2)
library(data.table)
library(magrittr)
library(tidyr)
library(dplyr)
library(datasets)
```
2. Load yeast data and required packages using the following code:
```{r}
gene <- fread("./extdata/eqtl/gene.txt")
genotype <- fread("./extdata/eqtl/genotype.txt")
genotype <- melt(genotype, id.vars = 'strain', variable.name = 'marker',
value.name = 'genotype')
growth <- fread("./extdata/eqtl/growth.txt")
growth <- melt(growth, id.vars = "strain", variable.name = 'media',
value.name = 'growth_rate')
marker <- fread("./extdata/eqtl/marker.txt")
```

# Section 01 - Warm Up: Choosing the right test
1. You are a data science consultant helping researchers pick the right tests to evaluate their hypotheses. For
each hypothesis, indicate which test from the ones you have seen in the lecture would be most appropriate:

a) A researcher collects data on the height (measured in cm) and weight (measured in g) of Germans.
She hypothesizes that there is a significant association between how tall Germans are and how much
they weigh. She would like to test this hypothesis without making any distributional assumptions.
spearman correlation

b) A researcher collects data on the weight (measured in g) of Bavarians before and after the Oktoberfest.
She would like to know whether there is a significant difference in average weight after the Oktoberfest
as compared to before it. Prior research indicates that the weight of Bavarians is approximately
normally distributed.
t-test (/welch or wilcoxon)

c) A researcher is evaluating a rapid antigen test. The company manufacturing the test claims that if
someone is infected, the test will correctly return a positive result 99% of the time. The researcher
hypothesizes that, in practice, the test is often improperly administered and therefore significantly less
sensitive. She asks 1000 individuals, which have all been confirmed to be infected by a PCR test, to
self-administer the antigen test. She records how often the antigen test correctly returns a positive
result.
binomial test p=0.99 as null hypothesis

d) The company manufacturing the test has collected a bigger dataset, comprising both infected and
non-infected individuals. For each individual, they record two datapoints: the result of a PCR test
(infected/not-infected), which is taken as ground truth, and the result of a self-administered antigen
test (positive/negative). They would like to show that, even if self-administered, the test still gives
some information about infection status and thus is better than nothing.
fischer exact test



Section 02 - Test the association between markers and growth
1. Reproduce the growth boxplot for marker 5211 using the following code:
```{r}
getMaltoseDt <- function(mrk){
growth_mrk <- merge(growth, genotype[marker == mrk, .(strain, genotype)],
by = 'strain')
growth_mrk[media == "YPMalt"]
}
# boxplot
plot_growth_one_mk <- function(mk){
ggplot(getMaltoseDt(mk), aes(genotype, growth_rate)) +
geom_boxplot() +
labs(title = mk) + theme_bw(base_size = 16) +
theme(plot.title = element_text(hjust = 0.5))
}
plot_growth_one_mk("mrk_5211")
```

2. Last week, using permutation, we saw that some markers associated with growth. Which of the statistical
tests from the lecture would you use to test this association? For each marker, we are not certain if the
genotype will cause a positive or negative effect on growth; therefore, which kind of alternative hypothesis
would you choose: double-sided, right or left? Apply the test to marker 5211 to obtain a p-value and see if
the association that we found last week still holds. If more than one of the tests are appropriate, use them
all and compare the results.
double-sided
```{r}
data <- getMaltoseDt("mrk_5211")

t.test(growth_rate ~ genotype, data = data, alternative="two.sided")
wilcox.test(growth_rate ~ genotype, data = data, alternative="two.sided")

```

3. Given a marker and the name of a statistical test, make a function that returns the p-value of the
association of that marker with respect to growth. You can use the template below to construct this
function. Then test your function on the markers 1653 and 5091. Since we used the same markers last week:
do you get similar results as last week?
```{r}
test_growth <- function(mk, test){
  m_dt <- getMaltoseDt(mk)
  if(test == 'wilcoxon') {
    pval <- t.test(growth_rate ~ genotype, data=m_dt, alternative="two.sided")$p.value
  } else {
    pval <- wilcox.test(growth_rate ~ genotype, data=m_dt, alternative="two.sided")$p.value  
  }
  return(pval)
}

test_growth("mrk_1653", 'wilcoxon')
test_growth("mrk_1653", '')
test_growth("mrk_5091", 'wilcoxon')
test_growth("mrk_5091", '')

```



# Section 03 - Correlations and correlation tests
1. Investigate the correlation between Sepal.Length and Sepal.Width within the iris dataset. Calculate
the correlation and plot your results. Are there any issues with your results? Discuss.
2. Repeat the previous analysis for each species independently. How does the correlation between
Sepal.Length and Sepal.Width change?

