---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r}
plot(cars)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
```{r}
n <- 1000
x <- seq(1,n)
sum(x)
```

```{r}
log(100, base=10)
```

```{r}
library(dslabs)
data(murders)

str(murders)

murders

a <- murders$region


```


```{r}
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")
temp <- c(35, 88, 42, 84, 81, 30)

names(temp) <- city

print(temp[1:3])
print(temp[c("Paris", "San Juan")])
```


```{r}
12:73

seq(1,100, 2)
```



```{r}
seq(6, 55, 4/7)
```


## Section 04

```{r}
library(dslabs)
data("murders")

pop <- murders$population

pop
```


```{r}
ranks <- rank(murders$population)
my_df <- data.frame(state=murders$state, ranks=ranks)

my_df


```

```{r}
temp <- c(35, 88, 42, 84, 81, 30)
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")

names(temp) <- city

temp_F <- 5/9 * (temp-32)


```


```{r}
data("na_example")

str(na_example)

ind <- is.na(na_example)

sum(as.integer(ind))

table(ind)

mean(na_example[!ind])

mean(na_example, na.rm=TRUE)
```


