# Section 00 - Getting ready
1. Make sure you have already installed and loaded the following libraries:
```{r}
library(ggplot2)
library(data.table)
library(magrittr) # Needed for %>% operator
library(tidyr)
```

# Section 01 - Color guidelines
What are best practices when using color for data visualizations? Select all that apply.
1. Avoid having too many colors for categorical data. x
2. Use color only when it actually adds meaning to the plot. x
3. Use divergent color scales for categorical data types.

# Section 02 - Correlation and Causation
Read the following statements. Decide for each statement separately whether it is true or false and give an
explanation.
1. The concept of reverse causality states that whenever A causes B, B also causes A. FALSE
2. If A causes B and A causes C, then B also causes C. FALSE
3. If A and B correlate and A happens before B, then A causes B. FALSE
4. Causation implies linear association. FALSE

# Section 03 - Effect of a third variable
Investigate the file coffee_sim.csv (simulated dataset) by first loading it as a data.table.
```{r}
coffee_dt <- fread("./extdata/coffee_sim.csv")
coffee_dt
summary(coffee_dt)
```
1. Suggest an appropriate visualization and implement it with ggplot2 to display a possible association
between coffee consumption and “datavizitis” disease risk, measured in deaths per 1000 individuals. Does
this plot by itself seem consistent with a causal effect of coffee on datavizitis?
```{r}
coffee_dt[,packs_cigarettes_per_day := as.factor(packs_cigarettes_per_day)]
coffee_dt[,coffee_cups_per_day := as.factor(coffee_cups_per_day)]

ggplot(coffee_dt, aes(coffee_cups_per_day, datavizitis_risk)) +
  geom_boxplot()

```
2. Investigate the full dataset. Do you see evidence for a third variable influencing association? Support your
statement with an appropriate plot. Draw a graph with the potential causal relationships you find consistent
with the data. Relate it to one of the situations from the lecture script’s figure 6.3 or Simpson’s paradox.
```{r}
ggplot(coffee_dt, aes(packs_cigarettes_per_day, datavizitis_risk)) +
  geom_boxplot()

ggplot(coffee_dt, aes(coffee_cups_per_day, datavizitis_risk, color=packs_cigarettes_per_day)) +
  geom_boxplot()


ggplot(coffee_dt, aes(packs_cigarettes_per_day, datavizitis_risk, color=coffee_cups_per_day)) +
  geom_boxplot()


ggplot(coffee_dt, aes(coffee_cups_per_day, datavizitis_risk)) +
  geom_boxplot() +
  facet_wrap(~packs_cigarettes_per_day)


ggplot(coffee_dt, aes(packs_cigarettes_per_day, datavizitis_risk)) +
  geom_boxplot() +
  facet_wrap(~coffee_cups_per_day)

```
common cause
smoking -> datavisitis
smoking -> coffee


# Section 04 - General guidelines in data visualization
Below is a graph taken from one published paper. Read the figure legend.
In biology, a wild type strain (WT) is a strain whose genome has not been artificially modified. In contrast,
a mutant is a strain whose genome has been artificially modified.
1. Discuss in groups what could be better representations. Use pen and paper. There can be many options.

- box plot
- different color for WT and Mutants
- color for below and above WT

2. Implement the solution proposed by the tutor.
```{r}
# simulate data
dt <- data.table(pro_uptake = c(rnorm(3, 10100, 300), rnorm(4, 12100, 300),
rnorm(3, 9850, 300), rnorm(4, 11100, 300),
rnorm(4,8300, 300), rnorm(3,10050, 300),
rnorm(3, 12000, 300), rnorm(3, 10020, 300),
rnorm(3, 10080, 300), rnorm(3, 10070, 300) ),
mutants = c(rep("WT",3), rep("T49A",4), rep("K227N",3), rep("A400V",4),
rep("L421P",4), rep("I500T",3), rep("N591D",3),
rep("A601T",3), rep("E684D",3), rep("G710R",3) ) )

dt
```
```{r}
means <- aggregate(pro_uptake ~ mutants, dt, mean)
means

dt[, mean := mean(pro_uptake), by=mutants]

wt_mean <- dt[mutants=="WT", mean][1]
wt_mean

dt[, color:=ifelse(mean == wt_mean, "grey", ifelse(mean > wt_mean, "red", "blue"))]
dt

#ggplot(dt, aes(mutants, pro_uptake)) +
#  stat_summary(fun=mean, colour="darkred", geom="point", 
#               shape=18, size=3, show.legend=FALSE) + 
#  geom_point()

level_order <- dt[order(mean), unique(mutants)]
level_order

ggplot(dt, aes(factor(mutants, level=level_order), pro_uptake, color=color)) +
  geom_boxplot() +
  geom_point(color="black") +
  xlab("mutant") +
  ylab("proline uptake")

```


# Section 05 - Covid19 fatality rate in Belgium
Consider the following dataset which contains the fatality rate (percentage of deaths among infected) by
COVID 19 of different age groups in Belgium in June 2020 (taken from De Smet, D. (2020). Is corona erger
dan de griep? De Standaard, 22 June 2020). The overall rate (all_ages) suggests that being a female
increases the risk of dying from COVID-19 upon infection. Using appropriate plots discuss the validity of
this hypothesis and draw a graph with the potential causal relationships you find consistent with the data.
Relate it to one of the situations from the lecture script’s figure 6.3 or Simpson’s paradox.
```{r}
fatality_dt <- fread('extdata/belgium_infection_fatality_rate_june2020.csv')
fatality_dt

fatality_long_dt <- melt(fatality_dt, id.vars = "age_group", variable.name = "gender", value.name = "death_rate")
fatality_long_dt

ggplot(fatality_long_dt[age_group!="All ages"], aes(age_group, death_rate, fill=gender)) +
  geom_col(position="dodge") +
  geom_text(aes(label=death_rate), position=position_dodge(width=0.9), vjust=-0.25) 
```
Simpsons paradox
There are more older women in Belgium, because older people are more susceptible to dying

two factors show certain association, when stratifying by third variable, The association changes direction

sex -> age
age -> death_rate
sex -> death_rate


# Section 06 - Smoking and Datavizitis severity
Consider the following dataset contains a population of 2000 individuals who got datavizitis. It consists of
the number of cigaretes each individual smokes per day, the severity of their datavizitis and if they were
hospitalized or not.
```{r}
datavizitis_smoking_dt <- fread("./extdata/datavizitis_smoking.csv")
datavizitis_smoking_dt
```
1. Visualize the relationship between the number of cigarettes smoked per day and datavizitis severity among
hospitalized individuals. Make use of geom_smooth(method="lm") to highlight the general trend.
```{r}
ggplot(datavizitis_smoking_dt[hospitalized=="Yes"], aes(cigarettes_per_day, datavizitis_severity)) +
  geom_point() +
  geom_smooth(method="lm")
```


2. Visualize the relationship between datavizitis severity and cigarettes smoked per day among all population.
Make use of geom_smooth(method="lm") to highlight the general trend.
```{r}
ggplot(datavizitis_smoking_dt, aes(cigarettes_per_day, datavizitis_severity)) +
  geom_point() +
  geom_smooth(method="lm")
```

3. Visualize the same relationship distinguishing between hospitalized and all individuals.
```{r}
ggplot(datavizitis_smoking_dt, aes(cigarettes_per_day, datavizitis_severity, color=hospitalized)) +
  geom_point() +
  geom_smooth(method="lm")

ggplot(datavizitis_smoking_dt, aes(cigarettes_per_day, datavizitis_severity)) +
  geom_point(aes(color=hospitalized)) +
  geom_smooth(aes(color=hospitalized), method="lm") +
  geom_smooth(method="lm")
```

4. Recent studies have looked at hospitalized patients who tested positive for Covid19 and their smoking
status. They propose smoking may provide a lower risk of developing severe Covid19 based on a negative
association between Covid19 severity and smoking status. Considering the previous results on datavizitis can
you come up with a different explanation? Draw a graph with the potential causal relationships you find
consistent with the data. Relate it to one of the situations from the lecture script’s figure 6.3 or Simpson’s
paradox.
```{r}
ggplot(datavizitis_smoking_dt, aes(hospitalized, cigarettes_per_day)) +
  geom_boxplot()

ggplot(datavizitis_smoking_dt, aes(hospitalized, datavizitis_severity)) +
  geom_boxplot()
```

because we subset by hospitalization. People are more likely to be hospitalized if they either smoke a lot 
or have a high severity of datavisitis (or both).
This can lead to a seemingly negative association between smoking and datavisitis when subsetting by hospitalization

common consequence
cigarettes -> hospitalization
datavisitis_sev -> hospitalization



# Section 07 - Supporting hypotheses with visualizations

1. Read the titanic.csv file into a data.table. You can read the description of the dataset on kaggle:
https://www.kaggle.com/c/titanic/data.
```{r}
titanic_dt <- fread("./extdata/titanic.csv")
titanic_dt
```

2. Inspect the data table and make a summary of the variables in the dataset. What is the overall passenger
survival rate?
```{r}
summary(titanic_dt)

titanic_dt[survived==1, .N] / titanic_dt[, .N]
titanic_dt[, mean(survived)]

```

3. Does age associate with survival? Make a plot showing the distribution of age per survival outcome.
```{r}
#titanic_dt[, .N , by=c("age", "survived")]

ggplot(titanic_dt, aes(age)) +
  geom_histogram() +
  facet_wrap(~survived)

ggplot(titanic_dt, aes(factor(survived), age)) +
  geom_violin()

ggplot(titanic_dt, aes(factor(survived), age)) +
  geom_boxplot() +
  geom_jitter(height=.1, weight=.1)
```

4. Visualize the relationship between passenger class and survival rate.
```{r}
ggplot(titanic_dt, aes(factor(survived), pclass)) +
  geom_jitter()

ggplot(titanic_dt[, .N, by=c("pclass", "survived")], aes(pclass, N, fill=factor(survived))) +
  geom_col(position = "dodge")

ggplot(titanic_dt[, .N, by=c("pclass", "survived")], aes(pclass, N, fill=factor(survived))) +
  geom_col(position = "fill")

ggplot(titanic_dt[, .(survival_rate = mean(survived)), by=pclass], aes(pclass, survival_rate)) +
  geom_col()


```

5. How is age distributed in each passenger class?
```{r}
ggplot(titanic_dt, aes(age, fill=factor(pclass))) +
  geom_histogram() +
  facet_wrap(~pclass)

ggplot(titanic_dt, aes(age, fill=factor(pclass))) +
  geom_histogram(aes(y=..density..)) +
  geom_density(alpha=0) +
  facet_wrap(~pclass)

ggplot(titanic_dt, aes(factor(pclass), age, color=factor(pclass))) + 
  geom_violin() +
  geom_boxplot(width=0.1)

ggplot(titanic_dt, aes(factor(pclass), age, color=factor(pclass))) + 
  geom_boxplot()
```


6. Considering the passenger class, do age and survival outcome associate? Given the findings on question 4,
comment on the results. Draw a graph with the potential causal relationships you find consistent with the
data. Relate it to one of the situations from the lecture script’s figure 6.3 or Simpson’s paradox.
```{r}
ggplot(titanic_dt, aes(factor(survived), age)) +
  geom_boxplot() +
  facet_wrap(~pclass)


```
indirect cause
age -> passenger class -> survival


