# Section 00 - Getting ready
1. Make sure you have already installed and loaded the following libraries:
```{r}
library(ggplot2)
library(data.table)
library(magrittr) # Needed for %>% operator
library(tidyr)
library(ggrepel)
```

# Section 01 - Plot types
Match each chart type with the relationship it shows best.
1. shows distribution and quantiles, especially useful when comparing uni-modal distributions.
boxplot

2. highlights individual values, supports comparison and can show rankings or deviations categories and totals
bar chart

3. shows overall changes and patterns, usually over intervals of time
line chart

4. shows relationship between two continuous variables.
scatterplot

Options: bar chart, line chart, scatterplot, boxplot

# Section 02 - Reproducing a plot with ggplot2
1. Reproduce the following visualization of the association between the variables cty and hwy for the years
1999 and 2008 from the dataset mpg using the library ggplot2:
```{r}
data(mpg)
mpg <- as.data.table(mpg)

mpg[, year:=as.factor(year)]

ggplot(data=mpg, aes(x=cty, y=hwy, color=year)) +
  geom_point() +
  geom_smooth(method="lm")
```

# Section 03 - Visualizing distributions

Iris is a classical dataset in machine learning literature. It was first introduced by R.A. Fisher in his 1936
paper. The dataset gives the measurements in centimeters of the variables sepal length and width and petal
length and width, respectively, for 50 flowers from each of 3 species of iris.

1. Load the iris data and transform it to a data.table. Have a look at its first and last rows.
```{r}
data(iris)
iris

head(iris)
tail(iris)
```


2. How are the lengths and widths of sepals and petals distributed? Make one plot of the distributions
with multiple facets. Hint: You will need to reshape your data so that the different measurements (petal
length, sepal length, etc.) are in one column and the values in another. Remember which is the best plot
for visualizing distributions.
```{r}
iris_long_dt <- melt(as.data.table(iris), id.vars = "Species", variable.name = "type", value.name = "length")
iris_long_dt

ggplot(data=iris_long_dt, aes(length)) + 
  geom_histogram() +
  facet_wrap(facets = "type")
```


3. Vary the number of bins in the created histogram. Describe what you see.
```{r}
# with too many its difficult to see the distributions
ggplot(data=iris_long_dt, aes(length)) + 
  geom_histogram(bins = 100) +
  facet_wrap(facets = "type")

# with too few we would miss bimodal distribution in petal width
ggplot(data=iris_long_dt, aes(length)) + 
  geom_histogram(bins = 5) +
  facet_wrap(facets = "type")
```

4. Visualize the lengths and widths of the sepals and petals from the iris data with boxplots.
```{r}
ggplot(data=iris_long_dt, aes(length)) + 
  geom_boxplot() +
  facet_wrap(facets = "type")

ggplot(data=iris_long_dt, aes(x=type, y=length)) +
  geom_boxplot()
```


5. Add individual data points as dots on the boxplots to visualize all points. Discuss: in this case, why is it
not good to visualize the data with boxplots? Hint: geom_jitter()

We miss the bimodal distributions in petal length and petal width
```{r}
ggplot(data=iris_long_dt, aes(x=type, y=length)) +
  geom_boxplot() +
  geom_jitter(width = .2, height = .2)
```


6. Alternatives to boxplot are violin plots (geom_violin()). Try combining a boxplot with a violinplot to
show the the lengths and widths of the sepals and petals from the iris data.
```{r}
ggplot(data=iris_long_dt, aes(x=type, y=length)) +
  geom_violin() +
  geom_boxplot(width=.05) 
```



7. Which pattern shows up when moving from boxplot to a violin plot? Investigate the dataset to explain
this kind of pattern, provide with visualization.
```{r}
iris_long_dt

ggplot(data=iris_long_dt, aes(x=Species, y=length, color=Species)) + 
  geom_jitter() +
  facet_wrap(facets = "type")

ggplot(data=iris_long_dt, aes(x=Species, y=length, color=Species)) + 
  geom_violin() +
  facet_wrap(facets = "type")

# This is the given solution (i think mine is better)
ggplot(data=iris_long_dt, aes(x=type, y=length, color = Species)) +
  geom_dotplot(binaxis="y", stackdir="centerwhole", dotsize=0.3)
```

# Section 04 - Visualizing relationships between continuous variables
1. Are there any relationships/correlations between petal length and width? How would you visually show it?
```{r}
iris

ggplot(data = iris, aes(x=Petal.Length, y=Petal.Width)) +
  geom_point() + 
  geom_smooth()

ggplot(data = iris, aes(x=Petal.Length, y=Petal.Width, color=Species)) +
  geom_point() + 
  geom_smooth()

```

2. Do petal lengths and widths correlate in every species? Show this with a plot.
```{r}
ggplot(data = iris, aes(x=Petal.Length, y=Petal.Width, color=Species)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap(facets = "Species", scales = "free")
```


# Section 05 - Axes scaling and text labeling

1. Load the medals dataset stored in the file medals.csv. Plot the total number of medals won against
population size in the 2016 Rio Olympics with a scatter plot. You can load the dataset with the following
code:
```{r}
medals_dt <- fread('extdata/medals.csv')
medals_dt

ggplot(data=medals_dt, aes(x=population, y=total, label=code)) +
  geom_point()

# x axis diffictult to read (need to log)
```

2. What are the problems with the previous plot? Solve these issues with an adapted version of the plot.
```{r}
ggplot(data=medals_dt, aes(x=population, y=total, label=code)) +
  geom_point() + 
  scale_x_log10() +
  scale_y_log10() +
  annotation_logticks()
```

3. Add the country labels to the points in the scatter plot. Compare the differences of using the library
ggplot2 and the library ggrepel for this task
```{r}
ggplot(data=medals_dt, aes(x=population, y=total, label=code)) +
  geom_point() + 
  scale_x_log10() +
  scale_y_log10() +
  annotation_logticks() +
  geom_text()

ggplot(data=medals_dt, aes(x=population, y=total, label=code)) +
  geom_point() + 
  scale_x_log10() +
  scale_y_log10() +
  annotation_logticks() +
  geom_text_repel()
```

# Section 06 - The importance of data visualization

Anscombe’s quartet was constructed in 1973 by the statistician Francis Anscombe to demonstrate both the
importance of graphing data before analyzing it, and the effect of outliers on statistical properties. anscombe
is directly built in R. You do not need to download it.

1. We reshaped the original anscombe data to anscombe_reshaped. Which one is tidier? (reshaped)
```{r}
anscombe_reshaped <- anscombe %>%
as.data.table %>%
.[, ID := seq(nrow(.))] %>%
melt(id.var=c("ID")) %>%
separate(variable, c('xy', "group"), sep=1) %>%
as.data.table %>%
dcast(... ~ xy) %>%
.[, group := paste0("dataset_", group)]

anscombe
anscombe_reshaped[]
```

2. Compute the mean and standard deviation of each variable for each group. What do you see?
```{r}
anscombe_reshaped[,.(x_mean=mean(x), x_std=sd(x), y_mean=mean(y), y_std=sd(y)), by=group]
```

3. For each dataset, what is the Pearson correlation between x and y? Hint: cor() and Wikipedia1 for
Pearson correlation.
```{r}
anscombe_reshaped[, .(pcc=cor(x,y)), by=group]
```


4. Only by computing statistics, we could conclude that all 4 datasets have the same data. Now, plot x and
y for each dataset and discuss.
```{r}
ggplot(data=anscombe_reshaped, aes(x,y, color=group)) +
  geom_point() +
  facet_wrap(facets = "group")
```


5. Consider now the datasets given in the file boxplots.csv. Load the data and visualize the different
datasets with a boxplot. What do you observe? What can you conclude?
```{r}
boxplots_dt <- fread("./extdata/boxplots.csv")
boxplots_dt

boxplots_long_dt <- melt(boxplots_dt, variable.name = "group", value.name = "value")
boxplots_long_dt

summary(boxplots_long_dt)

ggplot(boxplots_long_dt, aes(value, color=group)) +
  geom_boxplot() +
  facet_wrap(facets = "group")
```


6. Exchange the boxplots by violin plots in the previous exercise. What do you observe? Do you conclude
the same as you did when visualizing the datasets with the boxplots
```{r}
ggplot(boxplots_long_dt, aes(group, value, color=group)) +
  geom_violin()

ggplot(boxplots_long_dt, aes(group, value, color=group)) +
  geom_violin() +
  geom_point(alpha = 0.2, position=position_jitter(.3, .5))
```

# Section 07 - Understanding and recreating boxplots

1. Using the mtcars dataset, make a boxplot of the miles per gallon (mpg) per cylinder (cyl).
```{r}
mtcars_dt

mtcars_dt[,cyl:=as.factor(cyl)]

ggplot(mtcars_dt, aes(cyl, mpg)) +
  geom_boxplot() 
```

2. Now, recreate the same plot without using geom_boxplot. You have to add all the layers manually: IQR
box, median line, whiskers and outlier points. Hint: Remember how a boxplot is constructed2. You may
find these functions useful: IQR, geom_crossbar, geom_segment, geom_point. Use data.table commands.
```{r}
mtcars_dt[, c("q1", "median", "q3") := as.list(quantile(x=mpg, probs = c(.25, .5, .75))), by = cyl] 

mtcars_dt[, iqr := q3-q1]
mtcars_dt[, out_min := q1-iqr*1.5]
mtcars_dt[, out_max := q3+iqr*1.5]

mtcars_dt[, is_outlier:=(mpg < out_min | mpg > out_max)]

mtcars_dt
mtcars_dt

ggplot(mtcars_dt, aes(cyl, median, ymin=q1, ymax=q3)) +
  geom_crossbar() +
  geom_segment(aes(cyl, q3, xend = cyl, yend = out_max)) +
  geom_segment(aes(cyl, out_min, xend = cyl, yend = q1)) +
  geom_point(data=mtcars_dt[is_outlier==TRUE], aes(cyl, mpg))
```




