---
title: "R Notebook"
output: html_notebook
---

```{r}
library(data.table)
library(magrittr)
```
# Section 1
1
```{r}
ratings_dt <- fread("./extdata/Lecture02_extdata/Exercise_data/BX-Book-Ratings.csv")
books_dt <- fread("./extdata/Lecture02_extdata/Exercise_data/BX-Books.csv")
users_dt <- fread("./extdata/Lecture02_extdata/Exercise_data/BX-Users.csv")
```

3
```{r}
users_dt$Age <- as.integer(users_dt[, Age])
users_dt$Age[is.na(users_dt$Age)] <- 0

# Alternative
# users_dt[, Age := as.numeric(Age)]

```
4 5
```{r}
summary(books_dt)
sapply(users_dt, class)

head(ratings_dt, n=5)
tail(ratings_dt, n=5)
```
6
```{r}
colnames(ratings_dt) <- gsub('-', '_', colnames(ratings_dt))  
colnames(books_dt) <- gsub('-', '_', colnames(books_dt))  
colnames(users_dt) <- gsub('-', '_', colnames(users_dt))  
```

7
```{r}
books_dt[, c("Image_URL_S", "Image_URL_M", "Image_URL_L") := NULL]
```

8
Create a table book_dt_2 that contains all the books published between 1900 and 2019 (inclusive) from
the table books_dt.
```{r}
book_dt_2 <- books_dt[Year_Of_Publication >= 1900 & Year_Of_Publication <= 2019]
```

# Section 2
1
```{r}
length(unique(books_dt[, Book_Author]))

# Alternative
books_dt[, uniqueN(Book_Author)]
```

2
How many different authors are included for each year of publication between 2000 and 2010 (inclusive)
in books_dt?
```{r}
books_dt[Year_Of_Publication >= 2000 & Year_Of_Publication <= 2010, uniqueN(Book_Author), by=Year_Of_Publication]
```

3
In how many observations is the age information missing in the ratings table users_dt?
```{r}
users_dt[Age == 0, .N]
```

4
What is the maximum rating value in the ratings table?
```{r}
max(unique(ratings_dt[, Book_Rating]))

ratings_dt[, max(Book_Rating, rm.na=TRUE)]
```

5. What is the most common rating value larger than 0?
```{r}
ratings_dt[Book_Rating != 0, .N, by = Book_Rating][N==max(N)]
```

6. Which are the book identifiers (ISBN) with the highest ratings?
```{r}
ratings_dt[Book_Rating == 10, ISBN] %>% head

ratings_dt[Book_Rating == max(Book_Rating, rm.na=TRUE), ISBN] %>% head
```

7. Reorder the ratings table according to the rating value of each book in descending order. Hint: order()
```{r}
ratings_dt[order(Book_Rating, decreasing = FALSE)]

ratings_dt[order(-Book_Rating)]

# setorder(ratings_dt, -Book_Rating)
```

# Section 3
1. Add a new column called High_Rating to the data table ratings_dt. The column has an integer 1 for
all observations with a rating value higher than 7. Hint: ifelse()
```{r}
ratings_dt$High_Rating <- ifelse(ratings_dt[, Book_Rating] > 7, 1, 0)

ratings_dt[, High_Rating := ifelse(Book_Rating > 7, 1, 0)]
```

2. How many observations are considered to be a high ranking? What is the proportion of high ranked
observations among all observations?
```{r}
ratings_dt[High_Rating == 1, .N]
ratings_dt[High_Rating == 1, .N] / ratings_dt[, .N]

ratings_dt[, sum(High_Rating)]
ratings_dt[, sum(High_Rating)/.N]
```

3. Which users did not give any rating to any book? Filter these users out from users_dt. Hint: There’s no
need to merge users_dt with ratings_dt, we are simply interested in the users that are not in ratings_dt.
```{r}
raters <- ratings_dt[,User_ID]
raters
users_dt[! User_ID %in% raters]
```

4. What is the most common age of users who rated at least one book?
```{r}
users_dt[User_ID %in% raters & Age != 0, .N, by=Age]
users_dt[User_ID %in% raters & Age != 0, .N, by=Age][N==max(N)]
```

5. On average, how many books did a user rate?
```{r}
ratings_dt[, .N, by=User_ID][,sum(N)] / ratings_dt[, uniqueN(User_ID)]

ratings_dt[, .N, by=User_ID][, mean(N, na.rm=TRUE)]
```

6. What is the title of the first published book with the highest ranking?
```{r}
ratings_dt[Book_Rating == max(Book_Rating) & Year_Of_Publication == min(Year_Of_Publication), Book_Title] %>% head(1)
```

7. In which year was a book with the largest number of ratings last published?
```{r}
ratings_dt[, N := .N, by=ISBN][N==max(N), Year_Of_Publication][1]

ratings_dt[ISBN == ratings_dt[, .N, by=ISBN][N==max(N), ISBN] , Year_Of_Publication][1]
```

8. Add to the table ratings_dt the highest ranking that each book received as a new column called
Max_Book_Ranking.
```{r}
ratings_dt[, Max_Book_Ranking := max(Book_Rating), by=ISBN]
ratings_dt
```

9. Subset the ratings_dt ratings table to contain only books written by the following authors:
```{r}
authors <- c("Agatha Christie", "William Shakespeare", "Stephen King",
"Ann M. Martin", "Carolyn Keene", "Francine Pascal",
"Isaac Asimov", "Nora Roberts", "Barbara Cartland", "Charles Dickens")
authors

ratings_dt[Book_Author %in% authors]
```

10. How many ratings has each author from the previous exercise 9? What is their max and average ranking?
```{r}
ratings_dt[Book_Author %in% authors, .(COUNT = .N, 
                                       MAX_RANK = max(Book_Rating, rm.na=TRUE), 
                                       AVR_RANK = mean(Book_Rating, rm.na=TRUE)),  by=Book_Author]
```

# Section 4
```{r}
library(readxl)
```

1. Using the summer_olympic_medals.xlsx file, which athlete won most bronze medals? Hint read_excel()
from readxl package.
```{r}
summer_olympics <- read_excel("./extdata/Lecture02_extdata/Exercise_data/summer_olympic_medals.xlsx", sheet="ALL MEDALISTS")
summer_olympics <- as.data.table(summer_olympics)
```

```{r}
summer_olympics[Medal=="Bronze", .N, by=Athlete][N==max(N)]
```


2. Are the columns Gender and Event_gender consistent? Find inconsistent gender entries.
```{r}
summer_olympics[substr(Gender,1,1) != Event_gender & Event_gender != 'X']
```

3. Which country won most medals? Which country has the highest ratio of silver medals? Use the data in
the country summary sheet starting at row 147 of the summer_olympic_medals.xlsx file.
```{r}
summer_olympics[, .N, by="NOC"][N==max(N)]

summer_olympics[, TOTAL_M := .N, by="NOC"]
summer_olympics[Medal=="Bronze", BRONZE_M := .N, by="NOC"]
summer_olympics[Medal=="Silver", SILVER_M := .N, by="NOC"]
summer_olympics[Medal=="Gold", GOLD_M := .N, by="NOC"]

unique(summer_olympics[, SILVER_M/TOTAL_M, by="NOC"][!is.na(V1)])[V1==max(V1)]
```

```{r}
breakdown_dt <- as.data.table(read_excel("./extdata/Lecture02_extdata/Exercise_data/summer_olympic_medals.xlsx", sheet='BREAKDOWN', range="A3:G29"))
breakdown_dt

breakdown_dt[, S_Ratio := Silver/`Grand Total`, ]

breakdown_dt[S_Ratio==max(S_Ratio)]
```


4. Which countries did participate, but without winning medals? Assume, that all countries listed in the
IOC COUNTRY CODES sheet participated. Hint you can quick fix the column names with make.names
and find set differences with setdiff.
```{r}
country_codes_dt <- as.data.table(read_excel("./extdata/Lecture02_extdata/Exercise_data/summer_olympic_medals.xlsx", sheet='IOC COUNTRY CODES 1'))

all_codes <- country_codes_dt[,`Int Olympic Committee code`]
#all_codes

medal_codes <- summer_olympics[, unique(NOC)]
#medal_codes

diff <- setdiff(all_codes, medal_codes)
diff

country_codes_dt[`Int Olympic Committee code` %in% diff, Country...1]

```



